---
title: "Auditor General's Office - The Efficiency of Norway's Hospital System from 2010-2021"
output: html_notebook
---

1. Obtain relevant data from Statistics Norway and the Norwegian Directorate of Health, 
preferably using the API. 

#Statistics Norway API

```{r}
options(encoding="UTF-8")
library(httr)
library(rjstat) #For JSON
url <- "https://data.ssb.no/api/v0/no/table/10261"
# 
data.tmp <- '
{
  "query": [
    {
      "code": "Region",
      "selection": {
        "filter": "vs:Helseregion330b",
        "values": [
          "H12",
          "H03",
          "H04",
          "H05"
        ]
      }
    },
    {
      "code": "ContentsCode",
      "selection": {
        "filter": "item",
        "values": [
          "Heldognsopphold",
          "Dagbehandlinger"
        ]
      }
    }
  ],
  "response": {
    "format": "json-stat2"
  }
}
'
d.tmp <- POST(url , body = data.tmp, encode = "json", verbose())
# Henter ut innholdet fra d.tmp som tekst deretter bearbeides av fromJSONstat
ssb.df.dagbehandling <- fromJSONstat(content(d.tmp, "text"))
# Viser datasettet
```

#Health Directorate API example

```{r}
library(httr)
library(tinytex)
library(tidyverse)
library(janitor)
library(kableExtra)
library(rjstat)
url <- "https://api.helsedirektoratet.no/innhold/nki/kvalitetsindikatorer/"
key <- "1c50d76931ba48f69d177c18eaf3c6a8"

ds<-GET(url, 
        add_headers("Ocp-Apim-Subscription-Key" = key),
        add_headers("Cache-Control"= "no-cache")) %>% 
  content(as = "text") %>% # extracting the data
  jsonlite::fromJSON(flatten = TRUE) # parsing to dataframe


url2 <- "https://api.helsedirektoratet.no/innhold/nki/kvalitetsindikatorer/0003-0010-5/data?contentType=application/json"

dg<-GET(url2, 
        add_headers("Ocp-Apim-Subscription-Key" = key),
        add_headers("Cache-Control"= "no-cache")) %>% 
  content(as = "text") %>% # extracting the data
  jsonlite::fromJSON(flatten = TRUE)

dge<-dg$AttachmentDataRows

dge <-   clean_names(dge)

dge_only_norway <- dge %>%
  filter(parent_name=="Hele landet") %>%
  filter(period_type=="Årlig")
```


2. Carry out a descriptive analysis of key input factors, activity level and indicators for quality at 
the health region / hospital level. 

##Input factors
SSB
```{r}
##Pasienter på sykehus
url <- "https://data.ssb.no/api/v0/en/table/10261"
data.tmp <- '
{
  "query": [
    {
      "code": "Region",
      "selection": {
        "filter": "item",
        "values": [
          "0",
          "30",
          "01",
          "H12",
          "Uoppgitt"
        ]
      }
    },
    {
      "code": "Kjonn",
      "selection": {
        "filter": "item",
        "values": [
          "0",
          "1",
          "2"
        ]
      }
    },
    {
      "code": "Alder",
      "selection": {
        "filter": "item",
        "values": [
          "999",
          "00-09",
          "10-19",
          "70-79",
          "080+"
        ]
      }
    },
    {
      "code": "Diagnose3",
      "selection": {
        "filter": "item",
        "values": [
          "99",
          "02",
          "03",
          "92",
          "93"
        ]
      }
    },
    {
      "code": "ContentsCode",
      "selection": {
        "filter": "item",
        "values": [
          "PasientSomatSykeh",
          "Heldognsopphold",
          "Dagbehandlinger",
          "AntDagbehandlinger",
          "AntPoliklinKonsult"
        ]
      }
    },
    {
      "code": "Tid",
      "selection": {
        "filter": "item",
        "values": [
          "2012",
          "2013",
          "2014",
          "2020",
          "2021"
        ]
      }
    }
  ],
  "response": {
    "format": "json-stat2"
  }
}
'
d.tmp <- POST(url , body = data.tmp, encode = "json", verbose())
# Henter ut innholdet fra d.tmp som tekst deretter bearbeides av fromJSONstat
ssb.df.pasienter<- fromJSONstat(content(d.tmp, "text"))
# Viser datasettet
```


3. Visualize differences at the health region / hospital level, preferably with interactive figures 
in a report. 


4. Analyze / visualize the productivity of each hospital. Has productivity changed over time? 
How to compare productivity across hospitals? 


